---
title: "Predicting The Manner on How Barbell Exercise was Done"
output: html_document
---

Background

In this project, the goal was to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to predict the manner in which they did the exercise. This is the "classe" variable in the training set. The participants were asked to perform barbell lifts correctly and incorrectly in 5 different ways. This is the report describing how the model was built, how you cross validation was used, what the expected out of sample error is. This will also use the chosen prediction model to predict 20 different test cases. 


Data

The training data for this project are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har.

```{r, echo= FALSE}
library(caret)
library(RANN)
library(e1071)
library(klaR)
library(randomForest)
library(mboost)

set.seed(1)
fileUrl1 <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
if(!file.exists("training.csv")){
    download.file(fileUrl1,destfile="training.csv",method="curl")
}

dateDownloaded <- date()
dateDownloaded
data = read.csv("training.csv", na.strings = c("NA","","#DIV/0!"))


inTrain <- createDataPartition(y=data$classe,p=0.75, list=FALSE)
training <- data[inTrain,]
testing <- data[-inTrain,]


cleanTrain <- training[,8:159]

drops <- c("amplitude_yaw_belt", "amplitude_yaw_dumbbell", "amplitude_yaw_forearm",
           "kurtosis_yaw_belt", "skewness_yaw_belt", "kurtosis_yaw_dumbbell", "skewness_yaw_dumbbell", 
           "kurtosis_yaw_forearm", "skewness_yaw_forearm")
cleanTrain <- cleanTrain[,!(names(cleanTrain) %in% drops)]

cleanTest <- testing[,8:159]
cleanTest <- cleanTest[,!(names(cleanTest) %in% drops)]

cleanTrain <- cbind(cleanTrain, training$classe)
cleanTest <- cbind(cleanTest, testing$classe)

names(cleanTrain)[144] <- paste("classe")
names(cleanTest)[144] <- paste("classe")

imputeProc <- preProcess(cleanTrain[,-144],method="knnImpute")
imputedTrain <- predict(imputeProc, cleanTrain[,-144])

centerProc <- preProcess(log10(imputedTrain+1),method=c("center","scale"))
centeredTrain <- predict(centerProc, imputedTrain)

pcaProc <- preProcess(centeredTrain,method="pca", thresh = .95)
pcaTrain <- predict(pcaProc, centeredTrain)

imputedTest <- predict(imputeProc,cleanTest[,-144])
centeredTest <- predict(centerProc, imputedTest)
pcaTest <- predict(pcaProc, centeredTest)
```




Cleaning the Data


The data was cleaned with the following steps:
1. Field values with "NA" and "DIV\0" were condidred NA. 
2. Non-numeric variables and descriptive fields were removed from the dataset. 
3. The following variables with no data in all records were removed from the dataset: "amplitude_yaw_belt", "amplitude_yaw_dumbbell", "amplitude_yaw_forearm", "kurtosis_yaw_belt", "skewness_yaw_belt", "kurtosis_yaw_dumbbell", "skewness_yaw_dumbbell", "kurtosis_yaw_forearm", "skewness_yaw_forearm". 


Cross Validation 
1. 75% of the traning data were used as the training data set and the remaining 25% were to test the accuracy of the models tried. 

Pre-processing
1. The data were imputed to remove the NA values. The knnImpute method was used, with the default preProc arguments for this method. 
2. The data was also centered, using the "center" and "scale" methods, with the default preProc arguments for this methods. 
3. Lastly, PCA was applied to the resulting imputed, centerd and scaled data, with the default preProc arguments. 

The resulting training data that will be used is:

```{r, echo=TRUE}
pcaProc
```

Training

Four methods were used to train the data:
1. Naive Bayes
2. Random Forest
3. Generalized Linear Model
4. Lindear Discriminant Analysis

The resulting model from these methods were used to predict the classe for the testing data set (which was generated for cross validation out from the original training data set). Then a confusion matrix was run for the testing data set. 

The accuracy of the methods/models were used to determine goodness of fit of the models.


Random Forest generated the highest accuracy of 92% for the testing data set. This method and its generated model was used to predict the classe of the 20 problem records. 19 of the 20 observations were predicted correctly. 